# Tidy Evaluation & Metaprogramming

```{r, echo=FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  echo = TRUE
)
```

**Learning objectives:**

- Why data masking exists, and how it works, in theory
- Patterns for using data masking in practice

## Problem Part 1: reference {-}

- Data frame columns are not objects in the global environment...
- ... but we want to address them as such

```{r}
# proper search path: from environment object to column
mean(mtcars$cyl + mtcars$am)

# improper search: this is not the object you seek
mean(cyl + am)
```


## Problem Part 2: evaluation {-}

- Functions need to take data frame columns as arguments...
- ... but need to evaluate those names at the proper time and in proper context:     
    - upon execution of the function
    - in the data (frame) environment

```{r}
my_mean <- function(data, var1, var2) {
  dplyr::summarise(data, mean(var1 + var2))
}

my_mean(mtcars, cyl, am)
#> Error in `dplyr::summarise()`:
#> i In argument: `mean(var1 + var2)`.
#> Caused by error:
#> ! object 'cyl' not found
```


## How data masking solves problem 1 {-}

### Elevate the data environment  {-}

- Objects are defined in their environment
- Columns are defined in their data frame
- To enable users to refer to columns as objects, data-masking functions must:
  - Change the search path for column names
  - Give the data environment priority

### Disambiguate references {-}

- Suppose `col` is defined both in the global and data environment
- How does one refer to the right `col`?
- With pronouns
  - `.data` refers to column in the data environment
  - `.env` to an object in the global (parent) environment

## How defusal/injection solve problem 2 {-}

### Defuse {-}

- If bare column names are not defused, they will be evaluated immediately--hence in the wrong (global) environment
- If those names are defused, they need to be marked for later evaluation
- Mechanisms:
  - First quote (e.g., `enquo`) and evaluate (i.e., `!!`)
  - Embrace to quote and evaluate in one go for single columns: `{{` 

### Inject {-}

- If bare column names are provided as arguments, they must be "injected" into the function's body
- Mechanisms: same as above

### Summary of how data masking works {-}

1. **Defuse arguments.** First, quote bare names. Then, evaluate them later.
1. **Make data an environment.** Column names are resolved here. 
1. **Evaluate arguments explicitly.** 

## Data masking programming patterns : general problem {-}  

```
my_fun <- function(my_arg-my_behavior) {

  tidyverse::fun(my_arg-their_behavior)

}
```

## Data masking programming patterns {-}

- Argument behavior
- Patterns

## Argument behavior {-}

From the docs:

> In a regular function, arguments can be defined in terms of a type of objects that they accept.

> Data-masked arguments are more complex. Not only do they generally accept a specific type of objects ..., they exhibit special computational behaviours.

## Typology of argument behavior {-}

- **Data-masked expressions.** Expressions may refer to the columns of the supplied data frame.
- **Data-masked symbols.** Same as data-masked arguments but the supplied expressions must be simple column names.
- **Tidy selection.** alternative to data masking that supports selection helpers like starts_with(). There is in fact no masking at all. Expressions are either interpreted in the context of the data frame or evaluated in the user environment.
- **Dynamic dots.** These may be data-masked arguments, tidy selections, or just regular arguments. Dynamic dots support injection of multiple arguments

For more, read [here](https://rlang.r-lib.org/reference/topic-data-mask-programming.html#argument-behaviours)

## Argument behavior in the wild {-}

### Developers can signal argument behavior via roxygen2 tags... {-}

```
@param foo <[`data-masked`][dplyr::dplyr_data_masking]> What `foo` does.

@param bar <[`tidy-select`][dplyr::dplyr_tidy_select]> What `bar` does.

@param ... <[`dynamic-dots`][rlang::dyn-dots]> What these dots do.
```

### ... and they have {-}

- Data masking in [mutate](https://dplyr.tidyverse.org/reference/mutate.html).
- Tidy selection in, well, [select](https://dplyr.tidyverse.org/reference/select.html).
- Dynamic dots in [tibble](https://tibble.tidyverse.org/reference/tibble.html).

## Overview of data masking programming patterns {-}

1. **Forwarding.** Pass variables to another function. Inherit its behavior.
1. **Names.** Pass character vector of column names.
1. **Bridge.** Pass to bridge that accepts argument behavior, and from there pass to target function.
1. **Transformation.**

## Forwarding patterns {-}

> With the forwarding patterns, arguments inherit the behaviour of the data-masked arguments they are passed in.

With

- `{{`
- `...`

## Forwarding with `{{` {-}

Single argument

```{r}
# user-provided `var` is injected into the function body
my_summarise <- function(data, var) {
  data %>% dplyr::summarise({{ var }})
}
```


## Forwarding with `...` {-}

Multiple arguments

```{r}
# all user-provided elements of `...` are injected
my_group_by <- function(.data, ...) {
  .data %>% dplyr::group_by(...)
}

mtcars %>% my_group_by(cyl = cyl * 100, am)
```

## Names patterns {-}

>  function takes strings or character vectors of column names

Two patterns:

- One name
- Multiple names

## One name {-}

Use the `.data` pronoun

These are equivalent:

```{r}
# tidyverse
mtcars %>% dplyr::summarise(mean = mean(cyl))

# data masking-ish, with bare name
mtcars %>% dplyr::summarise(mean = mean(.data$cyl))

# data masking with character name <- THIS IS THE CODE OF INTEREST
var <- "cyl"
mtcars %>% dplyr::summarise(mean = mean(.data[[var]]))
```

## Multiple names {-}

- There is no plural of the `.data` pronoun (SNARKY ASIDE: but isn't data gramatically plural...?)
- But if target function accepts tidy select, use `all_of()`
- And if not, might need either
  - bridge pattern
  - symbolize-and-inject

Example of `all_of()`

```{r}
# NOTE: this is already a bridge pattern...
vars <- c("cyl", "am")
mtcars %>% tidyr::pivot_longer(
  # `all_of()` takes characters: https://tidyselect.r-lib.org/reference/all_of.html#ref-usage
  # but exhibits tidy-select behavior
  # which is what tidyr::pivot_longer expects for `cols`: https://tidyr.tidyverse.org/reference/pivot_longer.html#arguments
  cols = all_of(vars)
)
```

## Bridge patterns {-}

From the docs:

> Sometimes the function you are calling does not implement the behaviour you would like to give to the arguments of your function. [...] The general technique consists in forwarding the arguments inside a context that implements the behaviour that you want. Then, find a way to bridge the result to the target verb or function.

The common bridge patterns are:

- **selection -> data-mask.**
- **names -> data-mask.**
- **data-mask -> selection.**

## selection -> data-mask, via across() {-}

Setup:

- outer function, `my_group_by()`, expects selection
- inner function, `dplyr::group_by()`, [expects data-masking](https://dplyr.tidyverse.org/reference/group_by.html#ref-usage)

Building the bridge:

- across(), normally for mapping functions over columns, performs selections
-[ takes tidy selection in `.cols`](https://dplyr.tidyverse.org/reference/across.html#ref-usage)
- [returns tibble](https://dplyr.tidyverse.org/reference/across.html#value)

```{r}
# single selection
my_group_by <- function(data, var) {
  data %>% dplyr::group_by(across({{ var }}))
}

# multiple selections
my_group_by <- function(.data, ...) {
  .data %>% dplyr::group_by(
    across(  # <- expects tidy-select; returns tibble
      c(     # <- tidy-selection
        ...  # <- tidy-selection
      )
    )
  )
}
```

## names -> data-mask, via across(all_of()) {-}

Setup:

- outer function, `my_group_by()`, expects character names
- inner function, `dplyr::group_by()`, [expects data-masking](https://dplyr.tidyverse.org/reference/group_by.html#ref-usage)

Building the bridge:

- `all_of()` takes character names and is a tidy selector
- `across()` performs selection and returns a tibble

```{r}
my_group_by <- function(data, vars) {
  data %>% dplyr::group_by(
    across(        # <- takes selection; returns tibble
      all_of(vars) # <- takes character vector; performs tidy-selection
    )
  )
}

mtcars %>% my_group_by(c("cyl", "am"))
```

## data-mask -> selection, via transmute {-}

- `transmute()` creates a data frame that consists of columns in `...`
- `names()` yields a character vector of column names from `...`
- `all_of()` accepts a character vector of names and performs tidy selection

```{r}
my_pivot_longer <- function(data, ...) {
  # Forward `...` in data-mask context with `transmute()`
  # and save the inputs names
  inputs <- dplyr::transmute(data, ...)
  names <- names(inputs)
  
  # Update the data with the inputs
  data <- dplyr::mutate(data, !!!inputs)

  # Select the inputs by name with `all_of()`
  tidyr::pivot_longer(data, cols = all_of(names))
}

mtcars %>% my_pivot_longer(cyl, am = am * 100)
```

## Double Evaluation {-}

https://rlang.r-lib.org/reference/topic-double-evaluation.html

Tidy eval (and in particular, data-masking) introduces a bit of a problem that people 
may not have thought about before: 

> it accepts expressions that are ***computations***

We need to be careful about how we evaluate tidyeval expressions when we design 
functions, because we can fall into the trap of evaluating expressions more often
than we should. 

##### Avoid {-}

We can sometimes avoid double-evaluation issues by evaluating the expression to 
create an intermediate variable

(aside: lol transmute again?)

##### Symbols / glue strings {-}
Glue strings and other expression/symbol evaluation doesn't "suffer" from this problem

- "not suffering" meaning it doesn't execute code and instead translates the code to text

## Injection Operators Out Of Context {-}

https://rlang.r-lib.org/reference/topic-inject-out-of-context.html

Or, "what happens when you go overboard"

Injection operators `{{`, `!!`, and `!!!` are extensions of R syntax that only mean things
when evaluated in tidyverse context.

`{{` getting evaluated outside of tidy-eval is relatively harmless (except that it 
doesn't do what you want)

`!!` and `!!!` are negation operators and sometimes will cast things to logical 
if evaluated outside of tidyeval. It won't tell you that it did that either, so...
should avoid? (n+1 reasons to just use data masking instead of injections...)

## `{{` on regular objects {-}

https://rlang.r-lib.org/reference/topic-embrace-non-args.html

If you use `{{` on regular objects (and not on arguments), it captures the value 
of the object. This is related to how enquo works. 

## Function references {-}

https://rlang.r-lib.org/reference/eval_tidy.html - underlying function that powers data masking

https://rlang.r-lib.org/reference/dot-data.html - documentation for the pronouns

https://rlang.r-lib.org/reference/as_data_mask.html - underlying functions inside of eval_tidy


## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
