# Metaprogramming

```{r, echo=FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  echo = TRUE
)
library(rlang)
```

**Learning objectives:**

-   Explain why it's useful to know about defused expressions (maybe?).
-   Describe the relationship between defusing and lazy evaluation.
-   Compare and contrast the three types of defused expressions.
-   Differentiate between `expr()` (for defusing your own expressions) and `enquo()` & `enquos()` (for defusing expressions supplied by the user of your function).
-   Compare `{rlang}` defusing to base-R equivalents.

## Should we care about defused expressions? {-}

-   [`expr()` on CRAN](https://github.com/search?q=org%3Acran+%2F%28%5E%7C%5Cs%2B%7C%3A%29expr%5C%28%2F&type=code)
    -   Mostly literal metaprogramming? [{esquisse}](https://cran.r-project.org/package=esquisse), {gtsummary}, {tidyverse} packages
-   [`enquo()` on CRAN](https://github.com/search?q=org%3Acran+%2F%28%5E%7C%5Cs%2B%7C%3A%29enquo%5C%28%2F&type=code)
    - Mostly {tidyverse} & things interacting with it

Let's continue through to try to really grok this stuff, even if we* might never use them!

*Full disclosure: I will definitely use them/have used them: [{factory}](https://github.com/jonthegeek/factory/blob/07d9a555d3402f4dfc2abce9c42b886134e40abb/R/build_factory.R#L105) [{beekeeper}](https://jonthegeek.github.io/beekeeper/)

## Why "defuse"? {-}

```{r booby-trap}
ignore <- function(arg) NULL
force <- function(arg) arg

ignore(warning("boom"))
#> NULL

force(warning("boom"))
#> Warning in force(warning("boom")): boom

expr(force(warning("boom")))
#> force(warning("boom"))
```

> In a way, arguments are like booby traps which explode (evaluate) when touched. 
Defusing an argument can be seen as defusing the booby trap.

## What are the 3 types of defused expressions? {-}

| **type** | **examples** | **definition**                  |
|:---------|:-------------|:--------------------------------|
|calls     |`f(1, 2, 3)`<br />`1 + 1` |the action of calling a function to compute a new value|
|symbols   |`x`<br />`df`|named objects|
|constants |`1`<br />`NULL`|the actual object|

Note: Defused constants are still the constants themselves.

```{r defused-constants}
expr(1) == 1
#> TRUE
```

## How can we create defused expressions? {-}

`rlang::expr()` captures what you send in

```{r expr}
expr(foo)
#> foo

expr(mean(foo, na.rm = TRUE))
#> mean(foo, na.rm = TRUE)
```

## How else can we create defused expressions? {-}

Types ➡️ names of creation functions

```{r expr-types}
x <- "foo"
sym(x) == expr(foo)
#> TRUE

call("mean", sym(x), na.rm = TRUE) == expr(mean(foo, na.rm = TRUE))
#> TRUE
```

## Function references {-}

-   [expr()](https://rlang.r-lib.org/reference/expr.html)
-   [enquo() enquos()](https://rlang.r-lib.org/reference/enquo.html)
-   [base::eval()](https://rdrr.io/r/base/eval.html)
-   [eval_tidy()](https://rlang.r-lib.org/reference/eval_tidy.html)
-   [eval_bare()](https://rlang.r-lib.org/reference/eval_bare.html)
-   [quo_is_missing() quo_is_symbol() quo_is_call() quo_is_symbolic() quo_is_null() quo_get_expr() quo_get_env() quo_set_expr() quo_set_env()](https://rlang.r-lib.org/reference/quosure-tools.html)
-   [sym() syms() data_sym() data_syms()](https://rlang.r-lib.org/reference/sym.html)
-   [base::call()](https://rdrr.io/r/base/call.html)
-   [call2()](https://rlang.r-lib.org/reference/call2.html)
-   [base::bquote()](https://rdrr.io/r/base/bquote.html)
-   [base::substitute() base::quote() base::enquote()](https://rdrr.io/r/base/substitute.html)

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
