# Metaprogramming Part 2

```{r, echo=FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  echo = TRUE
)
library(rlang)
```

**Learning objectives:**

-   What is a quosure?
-   Why is a quosure?
-   How is a quosure?
-   Working with formulas

## What the heck is a quosure? {-}

"A quosure is a special type of defused expression that keeps track of the original context the expression was written in."

## Why do I need quosures? {-}

**User environment**:

```{r}
div100 <- function(x) {
  x / 100
}

dplyr::starwars %>%
  foo::summarise_bmi(mass, div100(height))
```

**Package foo**:

```{r}
bmi <- function(mass, height) {
  mass / height^2
}

summarise_bmi <- function(data, mass, height) {
  data %>%
    bar::summarise_stats(bmi({{ mass }}, {{ height }}))
}
```

**Package bar**:

```{r}
check_numeric <- function(x) {
  stopifnot(is.numeric(x))
  x
}

summarise_stats <- function(data, var) {
  data %>%
    dplyr::transmute(
      var = check_numeric({{ var }})
    ) %>%
    dplyr::summarise(
      mean = mean(var, na.rm = TRUE),
      sd = sd(var, na.rm = TRUE)
    )
}
```

**All add up to...**

```{r}
dplyr::transmute(
var =
  check_numeric(
    bmi(
      mass,
      div100(height)
    )
  )
)
```

With each tab over indicating a quosure boundary

- You need to make sure that `mass` and `height` stay properly tracked throughout environments

## Why? (In Practice) {-}

**Foreign and local expressions**:

```{r}
my_fun <- function(my_var) {
  their_function({{ my_var }})
  # or
  my_new_var <- enquo(my_var)
  their_function(!!my_new_var)
}
```

**External defusing with `...`**:

```{r}
my_group_by <- function(data, vars) {
  data |>
    dplyr::group_by(!!!vars)
}

my_group_by(data, dplyr::vars(X1, X2))
```

- `dplyr::vars()` creates quosures and passes them
  - An alias of `rlang::quos()`
- "As a rule of thumb, quosures are only needed for arguments defused with enquo() or enquos() (or with {{ which calls enquo() implicitly)"
  
## Okay, but what is a quosure? {-}

```{r, eval=TRUE}
rlang::quo(mtcars)

my_fun <- function(data) { print(rlang::quo(data)) }
my_fun(mtcars)
```

- A quosure has two parts:
  1. An expression (`rlang::quo_get_expr()`)
  2. An environment (`rlang::quo_get_env()`)
  
- A quosure is:
  - *callable*: Evaluation produces a result (currently require `rlang::eval_tidy()`)
  - *hygienic*: Evaluates in the tracked environment
  - *maskable*: "If evaluated in a data mask (currently only masks created with eval_tidy() or new_data_mask()), the mask comes first in scope before the quosure environment."
  
- Quosures are similar to promises for lazy evaluation, but quosures are repeatable and do not cache results
- Constants have an `empty` environment due to the structure of base R

A bit borrowed from Advanced R:

- Quosures are a subclass of formulas and are therefore call objects with an environment attribute

## Quosure Helpers {-}

- Defuse arguments as quosures: `enquo()` and `enquos()`
- Wrapping to create a new quosure: `quo()` and `quos()`
- Get the parts of a quosure: `quo_get_expr()` and `quo_get_env()`
- Assembling a new quosure: `new_quosure()` and `as_quosure()`

## Formulas {-}

- Formulas have a left hand side ~ and a right hand side or `~ "left" "right"`
- `{rlang}` provides helpers to 
  - access and set the left and right hand side
  - create new formulas with a specified environment

## Formula Helpers {-}

- Access and set the left and right hand side: `f_lhs()` and `f_rhs()`
- Access and set the environment: `f_env()`
- Create a new formula from parts: `new_formula()`
- Is this a formula?: `is_formula()` and `is_bare_formula()`

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
