# Conditions & Error Handling

**Learning objectives:**

-   Construct easy-to-understand error messages.
-   Use `{cli}` to format messages.
-   Bridge between `{rlang}` and `{cli}` messaging.
-   Customize `{cli}` output on your machine.
-   Recognize some additional condition functions.

## Brief intro to `cli_*()` {-}

`cli_abort()`, `cli_warn()`, and `cli_inform()` = preferred messaging

```{r cli-demo}
cli::cli_inform(c(
  "First unnamed formatted by handler type.",
  "v" = "Victory! (aka success).",
  "x" = "Danger.",
  "i" = "Info.",
  "!" = "Warning!"
))
```

## Error message content {-}

Detour: [tidyverse error messages](https://style.tidyverse.org/error-messages.html)

-   Tidyverse team has rules for error messages
-   `{cli}` was built to make messages that follow those rules
-   General idea:
    -   Short(ish), bulleted statements
    -   Sentence case + punctuation
-   Quick overview of rules on following slides

## Error message content: Problem statement {-}

> Every error message should start with a general statement of the problem. It should be concise, but informative (This is hard!).

-   Clear cause: use "must"
    -   "`n` must be a numeric vector, not a character vector."
    -   "`n` must have length 1, not length 2."
-   Else: use "can't"
    -   "Can't find column `b` in `.data`."
    -   "Can't coerce `.x` to a vector."
    -   "Can't find specified `.depth` in `.x`."
-   `cli`: Unnamed.

## Error message content: Error details {-}

1.    Describe exact error with `x = `.

  -   x = "Location 100 doesn't exist."
  -   x = "Result 1 is a character vector."
  -   If uncertain, list issues with `"*" =`
        -   c("\*" = "Size 2: Existing data.", "\*" = "Size 3: Column `y`.")

2.    Give contextual info with "i = ".

  -   i = "There are only 26 elements."
  -   i = "Only values of size one are recycled."

3.    If source of error is clear & common, provide a hint with `i = ... ?`
  -   i = "Did you mean `Species == \"setosa\"`?"
  -   i = "Did you use %>% instead of +?"

## Error message content: Grab bag {-}

More rules in [style guide](https://style.tidyverse.org/error-messages.html#punctuation), such as:

-   Don't capitalize arg/column name at start of bullet
-   Use singular in problem statements
-   If multiple problems, list up to 5, with "... and 5 more problems" (etc)
-   Aim for < 80 characters per bullet
-   Keep sentences simple for easier (eventual) localization/translation

## More about cli: Markup {-}

-   Can use `glue::glue()` syntax when `{glue}` is available
    -   `use_import_from("glue", "glue")`
-   Can also apply classes
    -   `adjective <- "great"; cli_text("This is {.emph {adjective}}.")`
-   Lots of classes (see [?cli::`inline-markup`](https://cli.r-lib.org/reference/inline-markup.html))
    -   `arg` = function argument,
    -   `cls` = S3/S4/R6/S7 class name
    -   `code` = code
    -   `help` = function help page, will link when possible
    -   Seriously see the [help](https://cli.r-lib.org/reference/inline-markup.html)!

## More about cli: Pluralization {-}

[Pluralization](https://cli.r-lib.org/articles/pluralization.html)

  -   "Found {nfile} file{?s}."
  -   "Found {ndir} director{?y/ies}."
  -   "Found {no(nfile)} file{?s}."
  -   "Will remove the {.pkg {pkgs}} package{?s}."
  -   "Will remove {?no/the/the} {.pkg {pkgs}} package{?s}."

## More about cli: Progress bars {-}

[Progress bars](https://cli.r-lib.org/articles/progress.html#the-current-progress-bar)

```{r progress-bar, eval = FALSE}
clean <- function() {
  wait_for_it <- function(.envir = rlang::caller_env()) {
    for (i in 1:100) {
      Sys.sleep(3/100)
      cli::cli_progress_update(.envir = .envir)
    }
  }
  cli::cli_progress_bar("Cleaning data #1", total = 100)
  wait_for_it()
  cli::cli_progress_bar("Cleaning data #2")
  wait_for_it()
}
clean()
```

## rlang to cli {-}

-   Be careful if you find/replace `rlang::abort()` to `cli::cli_abort()`
    -   Specifically, rewrite message-building to use `{cli}`
-   `on_load(local_use_cli())` to make `rlang::abort()` look `{cli}`-y
    -   `.onLoad <- function(lib, pkg) { run_on_load() }` in `aaa.R`

## Customizing cli {-}

-   Show characters for bullets: 
    -   `options(cli.condition_unicode_bullets = FALSE)`
-   Customize bullets: 
    -   `options(cli.user_theme = list(".cli_rlang .bullet-*" = list(before = "* "))`
-   Override colors: "background-color" & "color" in theme (see [vignette](https://rlang.r-lib.org/reference/topic-condition-customisation.html#changing-the-foreground-and-background-colour-of-error-calls))

## Leftovers {-}

[`cnd_header()`, `cnd_body()`, `cnd_footer()`](https://rlang.r-lib.org/reference/cnd_message.html)?

I think that's the only (maybe) relevant handler help we haven't talked about.

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
